`{peacesciencer}`: An R Package for Quantitative Peace Science Research

### 2021-11-10 user-guide/ms.Rmd
update

---
title: "`{peacesciencer}`: An R Package for Quantitative Peace Science Research"
thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
author:
- name: Steven V. Miller
  affiliation: Clemson University
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
# bibliography: inst/refs.bib
abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year, leader-year, leader-dyad-year, or state-year analysis (e.g. democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at `r params$ps_website`."
keywords: "software, statistical analysis, peace science, reproducibility"
geometry: margin=1in
csl: inst/jpr.csl
mainfont: cochineal
#sansfont: Linux Biolinum O 
fontsize: 11pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
  ps_website: ""
  ps_manuals: ""
  ps_github: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
#always_allow_html: TRUE
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \usepackage{csquotes}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
#  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
#  - \usepackage{colortbl}
#  - \arrayrulecolor{black}
#  - \usepackage{threeparttable}
#  - \usepackage{threeparttablex}
#  - \midrule\relax
---

@@

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists in the present. Afterward, it provides an overview of what is included in `{peacesciencer}` to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use `{peacesciencer}` in the R programming language. A more comprehensive tutorial follows, showing how `{peacesciencer}` already has a suite of data and functions that can allow for effective replications of a "dangerous dyads" type analysis [@bremer1992ddc] and standard state-year analyses of civil conflict onset [e.g. @fearonlaitin2003eicw]. This feature concludes with a discussion of what else `{peacesciencer}` is capable of doing in future updates and how this R package can inform more reasoned design decisions for researchers in peace science.

@@

`{peacesciencer}` is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch. NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars [@bennettetal2019n]. NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade [@bennettstam2000e]. No matter, `{peacesciencer}` is motivated by the following ideals that led to its creation.

For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g. dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. `{peacesciencer}` is primarily born from this question about how to create these data from scratch. The underlying code that produces these data types is available online and `{peacesciencer}` converts these lines of code into simple functions for the ease of the researcher.

@@

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer [c.f. @bowersvoors2016hiyr]. This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that `{peacesciencer}` uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet [@arelbundock2021w], scrape *any* type of data from the internet [@boehmke2016dwr], and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero [@arelbundock2021m]. `{peacesciencer}` embraces this. This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

@@

`{peacesciencer}` comes with a fully developed suite of built-in functions for generating some of the most widespread forms of peace science data and populating the data with important variables that recur in many quantitative analyses. The core functionality of `{peacesciencer}` reduces to two broad categories of functions. These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis. Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.

Table \@ref(tab:tab-createfunctions) are functions that crease base data frames for a researcher starting an original project. They serve as functions that communicate the units of analysis supported in this package and that the package is capable of generating for an interested user. For example, `create_stateyears()` will generate the full universe of state-years from the Correlates of War [@cowstates2016, v. 2016] or Gleditsch-Ward [@gledtischward1999rlis, v.2017] system, encompassing all state-years to the most recently concluded calendar year, depending on the arguments supplied to the user in the function. `create_leaderyears()` will generate the full universe of leader-years from the Archigos leader data [@goemansetal2009ia, v. 4.1], optionally standardizing leader-years to the Gleditsch-Ward or Correlates of War state system data. As of version 0.7, `{peacesciencer}` is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.[^leaderdyadyear] A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g. leader-months, state-quarters).

@@

```{r, echo=T, eval=FALSE}
install.packages(c("tidyverse", "peacesciencer"))
```

@@

```{r, echo=T, eval=FALSE}
library(tidyverse)
library(peacesciencer)
```

@@

```{r, eval=F, echo=T}
?create_stateyears()
```

This will open a documentation file for this function, which is itself quite verbose and informative for the user about what the function is doing. In this case, the documentation file will show there are three arguments in this function, each with built-in defaults. This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for CoW state system data or Gleditsch-Ward state system data and defaults to CoW), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years). If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2020, as of writing).

```{r}
create_stateyears()
```

@@

```{r}
create_stateyears(system = 'gw', subset_years = c(1946:2019))
```

@@

```{r, eval=F, echo=T}
?create_leaderyears()
```

@@

```{r}
create_leaderyears(standardize = "cow", subset_years = c(1870:2010))
```

@@

## Creating Dyad-Year Data and Adding Dyad-Year Data in `{peacesciencer}`

@@

```{r}
create_dyadyears(directed = FALSE, subset_years = c(1816:2010))
```

@@

## Creating State-Year Data and Adding State-Year Data in `{peacesciencer}`

@@

```{r, echo=T, eval=F}
create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
  filter(year %in% c(1946:2019)) %>%
  add_ucdp_acd(type="intrastate", only_wars = FALSE) %>%
  add_spells() %>%
  add_democracy() %>%
  add_creg_fractionalization() %>%
  add_sdp_gdp() %>%
  add_rugged_terrain() -> Data
```

@@

```{r, echo=T, eval=F}
create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
  add_ucdp_acd(type="intrastate", only_wars = TRUE) %>%
  add_spells() %>%
  rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
  left_join(Data, .) -> Data
```

@@

## Creating Leader-Year Data and Adding Leader-Year Data in `{peacesciencer}`

@@

Table \@ref(tab:tab-addfunctions) shows there are some dedicated functions for populating leader-level data with leader-specific information, in addition to adding state-year-level information (e.g. democracy, capabilities) to leader-level data.  Table \@ref(tab:tab-createfunctions) shows support for creating leader-level data that are standardized to either the CoW or G-W state system data. Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to @horowitzstam2014hpme. The user would first start with creating the base data (`create_leaderyears(standardize = "cow", subset_years=c(1875:2010))`), which also generates some leader attributes (e.g. estimated leader age, year in office, and leader gender). From there, they would add information about leader conflict behavior in the year with `add_gml_mids()` and calculate peace-years with the `add_spells()` function.[^oninitiation] `add_lead()` will add a battery of leader experience and attributes variables from data created by @ellisetal2015lead, including whether the leader had military experience, combat experience, was a rebel fighter, and many more. Finally, `add_nmc()` and `add_democracy()` will add state-year estimates of national capabilities and democracy for assisting in making state-to-state comparisons, even for leader-level analyses.

```{r, eval=F}
create_leaderyears(standardize = "cow", subset_years=c(1875:2010)) %>%
  add_gml_mids() %>%
  add_spells() %>%
  add_lead() %>%
  mutate(milservice_nocombat = ifelse(milservice == 1 & combat == 0, 1, 0)) %>%
  add_nmc() %>%
  add_democracy() -> Data

@@

@@

`{peacesciencer}` is already more than capable of creating the kind of data in high demand in peace science. It can create dyad-year, leader-year, leader-dyad-year, and state-year data (among others). It is also generalizable to the dispute data included in the package, allowing for merging into *dispute-year* data as well. This feature showed how it can effectively approximate two types of analyses in wide use in the peace science community. Surely researchers can and will add more information to these simple analyses after using `{peacesciencer}`, but the package already does a lot of the tedious work for researchers. It also does this in a maximally transparent way that conforms well to the DA-RT initiative across all political science. This is not to say `{peacesciencer}` does everything, but `{peacesciencer}` can only evolve and expand on what it already does well. This package can only evolve to meet new analytical demands for the peace science community. Users are free to request new features as "issues" on the project's Github.

>>>>

---
title: "`{peacesciencer}`: An R Package for Quantitative Peace Science Research"
thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
author:
- name: Steven V. Miller
  affiliation: Clemson University
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
# bibliography: inst/refs.bib
abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate common types of data from scratch and to merge in ubiquitous indicators included in almost every analysis (e.g. democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at `r params$ps_website`."
keywords: "software, statistical analysis, peace science, reproducibility"
geometry: margin=1in
csl: inst/jpr.csl
mainfont: cochineal
#sansfont: Linux Biolinum O 
fontsize: 11pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
  ps_website: ""
  ps_manuals: ""
  ps_github: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
#always_allow_html: TRUE
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \usepackage{csquotes}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
#  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
#  - \usepackage{colortbl}
#  - \arrayrulecolor{black}
#  - \usepackage{threeparttable}
#  - \usepackage{threeparttablex}
#  - \midrule\relax
---

@@

This data feature proceeds in the following fashion. The next section expands what need this package fills for peace scientists in the present. Afterward, it provides an overview of what is included in `{peacesciencer}` to help researchers more quickly conduct the kind of quantitative research they want. Thereafter, it provides a tutorial on how to install and best use `{peacesciencer}` in the R programming language. A more comprehensive tutorial follows, showing how `{peacesciencer}` already has a suite of data and functions that can allow for effective replications of a "dangerous dyads" type analysis [@bremer1992ddc], standard state-year analyses of civil conflict onset [e.g. @fearonlaitin2003eicw], and even leader-year analyses of inter-state dispute initiation [@horowitzstam2014hpme]. This feature concludes with a comparison of `{peacesciencer}` with other alternatives and a discussion of how `{peacesciencer}` can inform more reasoned design decisions for researchers in peace science.

@@

`{peacesciencer}` is motivated by the following ideals that led to its creation. For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g. dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. `{peacesciencer}` is primarily born from this question about how to create these data from scratch. The underlying code that produces these data types is available online and `{peacesciencer}` converts these lines of code into simple functions for the ease of the researcher.

@@

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer [c.f. @bowersvoors2016hiyr]. This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that `{peacesciencer}` uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet [@arelbundock2021w] and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero [@arelbundock2021m]. `{peacesciencer}` embraces this. This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

@@

`{peacesciencer}` comes with a fully developed suite of built-in functions for generating some of the most widespread forms of peace science data and populating the data with important variables that recur in many quantitative analyses. The core functionality of `{peacesciencer}` reduces to two broad categories of functions. These categories are functions that create the base data of interest to a researcher and functions, called after the base data are created, that add variables of interest to the data frame or subset the base data to a handful of rows that the researcher deems appropriate for analysis. Table \@ref(tab:tab-createfunctions) and Table \@ref(tab:tab-addfunctions) list these core functions as of version 0.7, the version of this package slated for release alongside the manuscript.[^helperfunctions]

[^helperfunctions]: `{peacesciencer}` also has miscellaneous "helper" functions that are not belabored in these manuscript. These include functions that allow the user to use `{peacesciencer}` functions with data created outside the package (`declare_attributes()`), download external data for quicker use of other functions (`download_extdata()`), and functions that communicate suggested citations (`ps_cite()`), among a few other functions. See the package's website for more information (`r params$ps_website`).

Table \@ref(tab:tab-createfunctions) are functions that create base data frames for a researcher starting an original project. They serve as functions that communicate the units of analysis supported in this package and that the package is capable of generating for an interested user. For example, `create_stateyears()` will generate the full universe of state-years from the Correlates of War [@cowstates2016, v. 2016] or Gleditsch-Ward [@gledtischward1999rlis, v.2017] system, encompassing all state-years to the most recently concluded calendar year, depending on the arguments supplied to the user in the function. `create_leaderyears()` will generate the full universe of leader-years from the Archigos leader data [@goemansetal2009ia, v. 4.1], optionally standardizing leader-years to the Gleditsch-Ward or Correlates of War state system data. As of version 0.7, `{peacesciencer}` is capable of creating full dyad-year data, leader-day data, leader-dyad-year data, leader-year data, state-day data, and state-year data.[^leaderdyadyear] A vignette on the package's website shows how users can create other forms of data from these functions as well (e.g. leader-months, state-quarters).

@@

```{r, echo=T, eval=FALSE}
# Install the packages for use
install.packages(c("tidyverse", "peacesciencer"))
```

@@

```{r, echo=T, eval=FALSE}
# Load the packages for use
library(tidyverse)
library(peacesciencer)
```

@@

```{r, eval=F, echo=T}
# See documentation for create_stateyears() function.
?create_stateyears()
```

This will open a documentation file for this function, which is itself quite verbose and informative for the user about what the function is doing. In this case, the documentation file will show there are three arguments in this function, each with built-in defaults. This function will allow the user to choose a state system for which they want state-years (`system`, which accepts either "cow" or "gw" for Correlates of War [CoW] state system data or Gleditsch-Ward [G-W] state system data and defaults to CoW), whether they want to extend the state system to the most recently concluded calendar year (`mry`, which defaults to TRUE), and whether they may want to additionally subset the years to a more narrow temporal domain they may already have in mind (`subset_years`, which defaults to no subset of the data, returning all possible state-years). If the user simply ran the function with no overrides, the function would return all Correlates of War state-years from 1816 to the most recent year (2020, as of writing).

```{r}
# Create all CoW state-years from 1816 to most recently concluded calendar year.
create_stateyears()
```

@@

```{r}
# Create all G-W state years from 1946 to 2019.
create_stateyears(system = 'gw', subset_years = c(1946:2019))
```

@@

```{r, eval=F, echo=T}
# See documentation for create_leaderyears() function.
?create_leaderyears()
```

@@

```{r}
# Create all leader-years for CoW states from 1870 to 2010.
create_leaderyears(standardize = "cow", subset_years = c(1870:2010))
```

@@

## Creating Dyad-Year Data and Adding to Dyad-Year Data in `{peacesciencer}`

@@

```{r}
# Create all non-directed dyad-years from 1816 to 2010.
create_dyadyears(directed = FALSE, subset_years = c(1816:2010))
```

@@

## Creating State-Year Data and Adding to State-Year Data in `{peacesciencer}`

@@

```{r, echo=T, eval=F}
# Create a set of state-year data for a civil conflict analysis.
create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
  filter(year %in% c(1946:2019)) %>%
  add_ucdp_acd(type="intrastate", only_wars = FALSE) %>%
  add_spells() %>%
  add_democracy() %>%
  add_creg_fractionalization() %>%
  add_sdp_gdp() %>%
  add_rugged_terrain() -> Data
```

@@

```{r, echo=T, eval=F}
# Repeat the process, but for a new DV of just wars
create_stateyears(system = 'gw', subset_years = c(1946:2019)) %>%
  add_ucdp_acd(type="intrastate", only_wars = TRUE) %>%
  add_spells() %>%
  rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
  left_join(Data, .) -> Data
```

@@

## Creating Leader-Year Data and Adding to Leader-Year Data in `{peacesciencer}`

@@

Table \@ref(tab:tab-addfunctions) shows there are some dedicated functions for populating leader-level data with leader-specific information, in addition to adding state-year-level information (e.g. democracy, capabilities) to leader-level data.  Table \@ref(tab:tab-createfunctions) shows support for creating leader-level data that are standardized to either the CoW or G-W state system data. Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to @horowitzstam2014hpme. The user would first start with creating the base data (`create_leaderyears(standardize = "cow", subset_years=c(1875:2010))`), which also generates some leader attributes (e.g. estimated leader age, year in office, and leader gender). From there, they would add information about leader conflict behavior in the year with `add_gml_mids()` and calculate peace-years with the `add_spells()` function.[^oninitiation] `add_lead()` will add a battery of leader experience and attributes variables from data created by @ellisetal2015lead, including whether the leader had military experience, combat experience, was a rebel fighter, and many more. Finally, `add_nmc()` and `add_democracy()` will add state-year estimates of national capabilities and democracy for assisting in making state-to-state comparisons, even for leader-level analyses. Some light recoding of the data created in `{peacesciencer}` and some regression modeling reproduces the results presented in Table \@ref(tab:tab-ly), itself an approximation of the kind of leader-year analysis exemplified in @horowitzstam2014hpme.

```{r, eval=F}
# Create a set of leader-year data for an analysis of conflict initiation.
create_leaderyears(standardize = "cow", subset_years=c(1875:2010)) %>%
  add_gml_mids() %>%
  add_spells() %>%
  add_lead() %>%
  add_nmc() %>%
  add_democracy() -> Data

@@

# A Comparison with Other Alternative Approaches

`{peacesciencer}` is not the only software available to peace science researchers who want to reduce the time and energy required to faithfully recreate data from scratch. Alternatives exist, some more inaccessible than others. NewGene, for example, is a stand-alone software program for Microsoft Windows and Mac that can create various types of data of interest to international relations scholars [@bennettetal2019n]. NewGene is itself the evolution of EUGene, which served conflict researchers well for over a decade [@bennettstam2000e]. Finally, peace scientists well-versed in Structured Query Language (SQL) could use one of several "join" transformations---e.g. "cross join", "inner join", "left join"---to create dyad-year data from state system data, even if this might amount to a detour in the peace scientist's research agenda for this particular task.[^joins] A comparison of `{peacesciencer}` with EUGene and NewGene suggests the following benefits of the package while emphasizing areas where other options may have some advantages.

[^joins]: That said, `{peacesciencer}` makes ample use of SQL-like "joins" for the construction of base data and functions that merge data into base data created in the package. A discussion of these joins appears as a vignette on the package's website.

EUGene is the clear inspiration for this package. Though its original impetus was the generation of expected utility data for evaluating @bdmlalman1992wr (i.e. the "EU" in "EUGene"), the software became quite popular for peace science scholars in the early 2000s for helping them start new projects from scratch with important data already provided. `{peacesciencer}` covers all the same units of analysis that EUGene covers, as of version 3.2.[^v3212] EUGene has more explicit support for dyadic dispute-year data whereas `{peacesciencer}` treats dyadic dispute-year data as a derivation of dyad-year data, with respect to functions that populate base data with additional information.[^whittledeclare] `{peacesciencer}` may reflect more current research interests in the peace science community. This is why there are leader-level data and support for the G-W (and UCDP) ecosystem of data that EUGene does not have, but there is no function yet for things like calculating Tau-b or expected utility values. Importantly, though, `{peacesciencer}` mimics the verbosity of EUGene's user manual. EUGene's user manual was amply informative about what it was calculating and why it was doing what it was doing for some defaults. Likewise, `{peacesciencer}`'s documentation strives to be as informative as possible as to what it is doing and why it is doing what it is doing for some defaults. A scholar who remembers EUGene well will ideally think of `{peacesciencer}` as the most faithful approximation of what that software did for the community at the time. It has the added benefit of being agnostic to the researcher's operating system, having greater flexibility of data types supported, and  better reflecting more current frontiers in the community. It does have the drawback of requiring at least some level of comfort with the R programming language.

[^whittledeclare]: For example, the "whittle" class of functions (i.e. `wc_` in Table \@ref(tab:tab-addfunctions)) are designed for dyadic dispute-year data while a helper function included in the package---`declare_attributes()`---can allow a user to treat dyadic dispute-year data as dyad-year data for the sake of using any of the other supported functions in Table \@ref(tab:tab-addfunctions).

[^v3212]: The latest version, as of writing, is v. 3.212 released in 2017. However, this is apparently a bug fix on string variables and that the underlying functionality was still last updated in 2007 with the release of version 3.2.

NewGene is the latest evolution of EUGene, at least as a stand-alone executable program for creating the kind of data of interest to the peace science community. NewGene's greatest strength, relative to EUGene and even `{peacesciencer}`, is its support for *k*-adic data [c.f. @poast2010mdd]. *k*-adic levels of analysis are not yet supported in `{peacesciencer}` and users interested in generating *k*-adic data should consider downloading NewGene. No matter, `{peacesciencer}` has several superlatives in relation to NewGene. It supports leader-year and leader-dyad-year data where NewGene does not. It offers support for the G-W (and UCDP) ecosystem of data whereas NewGene does not and is mostly aimed for researchers interested in inter-state conflict. NewGene deviates a bit from EUGene by only indirectly asking users what kind of data they want (e.g. state-year, dyad-year) and without providing too much detail about what it is doing and why it is doing what it is doing. For example, NewGene only indirectly states the unit of analysis of the data to be generated near the top of its interface by asking the user how many country columns they want. This is an indirect way of the user to get data that are state-year, dyad-year, triad-year (etc.), which are then expanded and populated with data at various levels. `{peacesciencer}`, much like EUGene, is more explicit, encouraging the user to be upfront about what their unit of analysis is and what are the units of analysis that can be plausibly plugged into the data the user is creating. `{peacesciencer}`, again, does expect at least some level of comfort with the R programming language, but even this comes with greater ease of interpreting what is the unit of analysis and what are the primary spatial and temporal units that serve as the basis of the data.

@@

`{peacesciencer}` is already more than capable of creating the kind of data in high demand in peace science. It can create dyad-year, leader-year, leader-dyad-year, and state-year data (among others). It is also generalizable to the dispute data included in the package, allowing for merging into *dispute-year* data as well. This feature showed how it can effectively approximate three types of analyses in wide use in the peace science community. Surely researchers can and will add more information to these simple analyses after using `{peacesciencer}`, but the package already does a lot of the tedious work for researchers. It also does this in a maximally transparent way that conforms well to the DA-RT initiative across all political science. This is not to say `{peacesciencer}` does everything, but `{peacesciencer}` can only evolve and expand on what it already does well. This package can only evolve to meet new analytical demands for the peace science community. Users are free to request new features as "issues" on the project's Github.
